### 2022-09-22 chapters/benchmarking/benchmarking.tex
refactoring and fix missing refrences
update :benchmarking protocol

\newpage
\chapter{A Benchmarking Protocol to Measure Software Energy Consumption}\label{chapter:benchmarking}

\import{\currfiledir}{draft}
\import{\currfiledir}{reproducibility}
\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

>>>>

\chapter{A Benchmarking Protocol to Measure Software Energy Consumption}\label{chapter:benchmarking}
\section{Introduction \note{missing} }
This chapter covers ways to overcome empirical analysis challenges in energy consumption studies. First, we go through the three components of a successful benchmark within the energy consumption experiment. Section~\ref{sec:benchmarking_reproducibility} concentrates on the "reproducibility" issue and how to deliver reproducible experiments without interfering with the energy measurement,while Section~\ref{sec:taming-the-energy-variation} discusses accuracy in software energy consumption.
\import{\currfiledir}{draft}
\import{\currfiledir}{experiment}
\import{\currfiledir}{conclusion}

### 2022-09-22 chapters/benchmarking/conclusion.tex
comments
update :benchmarking protocol

%% maybe add extension as a part of the conclusion for this chapter 
\section{Conclusion \note{Uncomplete }}
% TODO : uncomplete 
By the end of this study, we have gathered enough guidelines to make the tests more reproducible, accurate and extensible.
We created a set of new tests named \textbf{energy tests} which are more similar to performance tests.
Thanks to the work of two interns [Mamadou and Adrien] we created a CI/CD platform to measure the energy consumption of Java projects and we could track the evolution of the this energy across different stages of the project.
In the figure below we see an example of this plugin.
For more details please visit the gitlab repository\footnote{\url{https://gitlab.inria.fr/mamdiall/J-Joules}}~\footnote{\url{https://gitlab.inria.fr/mamdiall/jjoules-plugin}} .

We propose to extend the benchmarking framework suggested by the \emph{Collaboratory on Experimental Evaluation of Software and Systems in Computer Science}\footnote{\url{http://evaluate.inf.usi.ch/}} to address those limitations.
Instead of only presenting the four primary aspects of their guidelines that were mentioned previously in Section~\ref{sec:benchmarking_reproducibility}, we propose an abstract model to describe an empirical experiment:

\emph{context}: the hardware and the software configuration for the actual experiment, the purpose of this part is to provide extra information to help readers have a better judgement on the results and to be able reproduce the experiment while diminishing the impact of the external factor .
An \emph{orchestrator} : which its job to run the experiment.
The orchestrator provides three interfaces:

\emph{Workload interface}: provides a set of functions that should be implemented by the workload and called by each candidate in the experiment.
The Observer interface provides a set of metrics that are collected by the orchestrator during the experiment. Note that some of those metrics are external, such as the energy consumption of the system , execution time, execution stack, etc .
\begin{figure}[!htb]
    \center{\includegraphics[width=\linewidth]{imgs/benchmarkingprotocol}}
    \caption{Benchmarking protocol}\label{fig:benchmarkingprotocol}
\end{figure}
Figure~\ref{fig:benchmarkingprotocol} explains in detail the relationship within each component of the protocol.
First of all, we should start with the orchestrator which is the kernel of the experiment. Most of the time this kernel should represent the constants parts of the experiment.
The orchestrator is responsible for the following  for the coordination between the workload and the candidate, therefore it provides an interface that is implemented by each workload.
Later this interface will be provided to each candidate by the orchestrator. Like this we will remove any bias of dependency between the workload and the candidates, which will alow us to extend both parts without any change in the code.
Instead of retrieving the data directly from the experiment, we suggest to use an observator to allow the extension of the metrics, this observer in general should be outside the experiment to be able to gather as many informations without impacting the the code.

>>>>

%% maybe add extension as a part of the conclusion for this chapter 
\section{Conclusion \note{Uncomplete }}
In this chapter we :
- studied two aspects of the three creterias of the emperical research and how they will be applied in the field of energy measurement .
- we provided a brief comparaison of different methods to increase the reproducibility of the measurement .
- we studied the factors that can affect the measurement accuracy and how to reduce them .
- and finally

%  THIS SHOULD BE WITHIN THE perspective chapter 
% By the end of this study, we have gathered enough guidelines to make the tests more reproducible, accurate and extensible.
% We created a set of new tests named \textbf{energy tests} which are more similar to performance tests.
% Thanks to the work of two interns [Mamadou and Adrien] we created a CI/CD platform to measure the energy consumption of Java projects and we could track the evolution of the this energy across different stages of the project.
% In the figure below we see an example of this plugin.
% For more details please visit the gitlab repository\footnote{\url{https://gitlab.inria.fr/mamdiall/J-Joules}}~\footnote{\url{https://gitlab.inria.fr/mamdiall/jjoules-plugin}} .

### 2022-09-22 chapters/benchmarking/draft.tex
update :benchmarking protocol

%  NOTE : add introduction and change this into motivation ?? 
Empirical measurements are critical to capture the effect of developers' choices on software energy consumption.
To accomplish this, one should not overlook the benchmarking pitfalls highlighted in the state of the art~\cite{van_der_kouwe_benchmarking_2018}.
Furthermore, when it comes to evaluating the energy consumption of software, other challenges occur, such as the impact of hardware, the difference between the machines used for measurements and those that may apply the results, and so on.
Moreover, considerable advances in computer science have resulted in an increase in the number of obsolete results.
Addressing all of them becomes significantly more difficult in studies that compare multiple treatments.
Also, between the exploratory experiment and when the results were made public, new candidates may have come up and others may have changed.
So, not only is it important to make sure the results can be "reproduced," but it's also important to make room for new candidates.
This chapter covers ways to overcome empirical analysis challenges in energy consumption studies. First, we go through the three components of a successful benchmark within the energy consumption experiment. Section~\ref{sec:benchmarking_reproducibility} concentrates on the "reproducibility" issue and how to deliver reproducible experiments without interfering with the energy measurement. Section~\ref{sec:taming-the-energy-variation} discusses accuracy in software energy consumption.
Finally, Section~\ref{sec:bench_extension} presents tools for introducing new candidates, workloads, or other key performance indicators.
% Regarding the "representativeness" of the benchmarks, we will elaborate in the section entitled Perspectives.
%% this is might be the part that treats reproducibility for my case

@@

\subsection{Conclusion}
As can be seen, the introduction of the supervisor has increased the impact of Docker on energy that will be applied equitably to all experiments.
Therefore when it comes to comparison analysis, it will mitigate its impact automatically.
Furthermore, because we have access to the host Hardware, we do not need to worry about capturing the SUT's energy use.
As a result, we will use Docker as an encapsulation solution for all of our experiments.
It will ensure the reproducibility of our experiments in a simple and straightforward manner.
\clearpage

>>>>

%  NOTE : add introduction and change this into motivation ?? 
Empirical measurements are critical to capture the effect of developers' choices on software energy consumption.
To accomplish this, one should not overlook the benchmarking pitfalls highlighted in the state of the art~\cite{van_der_kouwe_benchmarking_2018}.
Furthermore, when it comes to evaluating the energy consumption of software, other challenges occur, such as the impact of hardware, the difference between the machines used for measurements and those that may apply the results, and so on.
Moreover, considerable advances in computer science have resulted in an increase in the number of obsolete results.
Addressing all of them becomes significantly more difficult in studies that compare multiple treatments.
Also, between the exploratory experiment and when the results were made public, new candidates may have come up and others may have changed.
So, not only is it important to make sure the results can be "reproduced," but it's also important to make room for new candidates.
% TODO :  something missing here  to link between these two paragraphs.

% Regarding the "representativeness" of the benchmarks, we will elaborate in the section entitled Perspectives.
%% this is might be the part that treats reproducibility for my case

@@

As can be seen, the introduction of the supervisor has increased the impact of Docker on energy that will be applied equitably to all experiments.
Therefore, when it comes to comparison analysis, it will mitigate its impact automatically.
Furthermore, because we have access to the host hardware, we do not need to worry about capturing the SUT's energy use.
As a result, we will use Docker to keep all of our tests distinct, as it will assure reproducibility in a simple and clear manner.

\subsection{Extension}
Whith a rapid evolution of the software industry, it is important to be able to include new candidates
We propose to extend the benchmarking framework suggested by the \em{Collaboratory on Experimental Evaluation of Software and Systems in Computer Science}\footnote{\url{http://evaluate.inf.usi.ch/}} to address those limitations.
Instead of only presenting the four primary aspects of their guidelines that were mentioned previously in Section~\ref{sec:benchmarking_reproducibility}, we propose an abstract model to describe an empirical experiment:

\em{context}: the hardware and the software configuration for the actual experiment, the purpose of this part is to provide extra information to help readers have a better judgement on the results and to be able reproduce the experiment while diminishing the impact of the external factor .
An \em{orchestrator} : which its job to run the experiment.
The orchestrator provides three interfaces:

\em{Workload interface}: provides a set of functions that should be implemented by the workload and called by each candidate in the experiment.
The Observer interface provides a set of metrics that are collected by the orchestrator during the experiment. Note that some of those metrics are external, such as the energy consumption of the system , execution time, execution stack, etc .
\begin{figure}[!htb]
    \center{\includegraphics[width=\linewidth]{imgs/benchmarkingprotocol}}
    \caption{Benchmarking protocol}\label{fig:benchmarkingprotocol}
\end{figure}

Figure~\ref{fig:benchmarkingprotocol} explains in detail the relationship within each component of the protocol.
First of all, we should start with the orchestrator which is the kernel of the experiment. Most of the time this kernel should represent the constants parts of the experiment.
The orchestrator is responsible for the following  for the coordination between the workload and the candidate, therefore it provides an interface that is implemented by each workload.
Later this interface will be provided to each candidate by the orchestrator. Like this we will remove any bias of dependency between the workload and the candidates, which will alow us to extend both parts without any change in the code.
Instead of retrieving the data directly from the experiment, we suggest to use an observator to allow the extension of the metrics, this observer in general should be outside the experiment to be able to gather as many informations without impacting the the code.

\subsection{Conclusion}

In this section, we have discussed the different options for encapsulating the system-under-test in order to ensure experiment reproducibility.
We have shown that the use of VMs is not suitable for performance-related tests, while containers are a good alternative. and then we discussed the pros and cons of using Docker while measuring the energy consumption of our experiments.

\clearpage

### 2022-09-22 chapters/benchmarking/experiment.tex
update :benchmarking protocol

\newpage
% To satisfy those three requirements. we elaborated a benchmarking protocol that aims to help reaserchers and practicioners in their experementations.
% for each aspect we present a set of guidelines and tools.
%% This is a conclusion maaan  !! 
\section{Improving Accuracy by Taming Energy Variations}\label{sec:taming-the-energy-variation}
\subsection{introduction}
While the previous section aimed to ensure the reproducibility of our software energy consumption experiments, this section provides a collection of tips and tools to help increase the accuracy of these experiments.
We are well aware of the effect that hardware has on energy fluctuations.
However, we feel that there is still an opportunity for practitioners to minimize this energy variance by employing solely tuneable factors.
To that end, we conducted a series of empirical studies utilizing state-of-the-art recommendations to discover which controllable elements can limit the variations of benchmark energy usage.
% We will first start with evaluating the parameters mentioned by \cite{heinrich_predicting}.
\subsection{Research questions}
This study will focus on the following research questions:
\begin{description}
    \item[\textsc{RQ}~1:] Does the benchmarking protocol affect the energy variation?
    \item[\textsc{RQ}~2:] How important is the impact of the processor features on the energy variation?
    \item[\textsc{RQ}~3:] What effect does the operating system have on energy variation?
    \item[\textsc{RQ}~4:] Does the choice of processor make a difference in reducing the energy variation?
\end{description}

@@

We used the platform Grid5000 (G5K)~\cite{grid5000,margery:hal-00965708}, which is a large, flexible testbed for experiment-driven research that is spread out across all of France, to meet these needs. Grid5000 has several clusters that are made up of 4 to 124 identical machines with different configurations for each cluster. For our experiment, we looked at four groups. Our main criterion was the CPU configuration.
The table~\ref{table:g5k}  below, presents a description of the four clusters that have been chosen for our experiments.

@@

\subsection{Conclusion}
In this part, we conducted an empirical study of controllable factors that can increase the energy variations on platforms with some of the latest CPU, and for several workloads.
We provide a set of guidelines that can be implemented and tuned (through the OS GRUB for example), especially with the new data centers isolation trend and the cloud usage, even for scientific and R\&D purposes.
Our guidelines aim at helping the user in reducing the CPU energy variation during systems benchmarking, and conduct more stable experiments when the variation is critical.
For example, when comparing the energy consumption of two versions of an algorithm or a software system, where the difference can be tight and need to be measured accurately.

Overall, our results are not intended to nullify the variability of the CPU, as some of this variability is related to the chip manufacturing process and its thermal behavior.
The aim of our work is to be able to tame and mitigate this variability along controlled experiments.
We studied some previously discussed aspects on some recent CPU, considered new factors that have not been deeply analyzed to the best of our knowledge, and constituted a set of guidelines to achieve the variability mitigating purpose.
Some of these factors, like the C-states usage, can reduce the energy variation up to 500\,\% at low workloads, while choosing the wrong cores/PU strategy can cause up to $30\times$ more variability.

We believe that our approach can also be used to study/discover other potential variability factors, and extend our results to alternative CPU generations/brands.
Most importantly, this should motivate future works on creating a better knowledge on the variability due to CPU manufacturing process and other factors.

>>>>

% To satisfy those three requirements. we elaborated a benchmarking protocol that aims to help reaserchers and practicioners in their experementations.
% for each aspect we present a set of guidelines and tools.
%% This is a conclusion maaan  !! 
\section{Improving Accuracy by Taming Energy Variations}\label{sec:taming-the-energy-variation}
\subsection{introduction}
While the previous section aimed to ensure the reproducibility of our software energy consumption experiments, this section provides a collection of tips and tools to help increase the accuracy of these experiments.
We are well aware of the effect that hardware has on energy fluctuations.
However, we feel that there is still an opportunity for practitioners to minimize this energy variance by employing solely tuneable factors.
To that end, we conducted a series of empirical studies utilizing state-of-the-art recommendations to discover which controllable elements can limit the variations of benchmark energy usage.
% We will first start with evaluating the parameters mentioned by \cite{heinrich_predicting}.
\subsection{Research questions}
This study will focus on the following research questions:
\begin{description}
    \item[\textsc{RQ}~1:] Does the benchmarking protocol affect the energy variation?
    \item[\textsc{RQ}~2:] How important is the impact of the processor features on the energy variation?
    \item[\textsc{RQ}~3:] What effect does the operating system have on energy variation?
    \item[\textsc{RQ}~4:] Does the choice of processor make a difference in reducing the energy variation?
\end{description}

@@

We used the platform Grid5000 (G5K)~\cite{grid5000,margery:hal-00965708}, which is a large, flexible testbed for experiment-driven research that is spread out across all of France, to meet these needs. Grid5000 has several clusters that are made up of 4 to 124 identical machines with different configurations for each cluster. For our experiment, we looked at four groups. Our main criterion was the CPU configuration.
Table~\ref{table:g5k}  below, presents a description of the four clusters that have been chosen for our experiments.

@@

\subsection{Conclusion}
In order to increase the accuracy in comparative experiments, we conducted an empirical study of controllable factors that may increase the energy variations on platforms with some of the latest CPUs and for several workloads.
this study is not intended to nullify the variability of the CPU, as some of this variability is related to the chip manufacturing process and its thermal behavior. Instead, it aims to tame and mitigate this variability along controlled experiments.
In this study, We investigated some previously discussed aspects on some recent CPU, considered new factors that have not been deeply analyzed to the best of our knowledge, and constituted a set of guidelines to increase the accuracy of energy measurement.
Some of these factors, like the C-states usage, can reduce the energy variation up to 500\,\% at low workloads, while choosing the wrong cores/PU strategy can cause up to $30\times$ more variability.
